---
title: "data_manipulation_dplyr"
author: "Jiayi Shi"
date: "2022-09-22"
output: github_document
---
```{r echo=FALSE,message=FALSE}
library(tidyverse)
```

## Load dataset
```{r}
litters_data = read_csv('data/FAS_litters.csv')
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```

## Select
```{r}
select(litters_data,1:5)
#select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
#select(litters_data,group:gd_of_birth)
#select(litters_data,group,gd_of_birth)
#specify columns you’d like to remove:select(litters_data, -group,-pups_survive) 
```

* Rename(won't change the original dataset, unless litters_data=...)
```{r}
#select(df, New_name = initial_name)#method1
select(litters_data,GROUP = group)
rename(litters_data, GROUP = group)#method2
```

* `starts_with()`, `ends_with()`, and `contains()`    
Useful functions in select, especially when variables are named with suffixes or other standard patterns.
```{r}
select(litters_data, starts_with("gd"))
select(litters_data, ends_with("weight"))
select(litters_data, contains("pups"))
```

*`everything()`,`relocate()``   
Pulling first (reorganizing columns)
```{r}
select(litters_data, litter_number, pups_survive, everything())#method1
relocate(litters_data, litter_number, pups_survive)#method2
```

## filter
`filter(df,condition)`  
`gd_of_birth == 20`  
`pups_born_alive >= 2`  
`pups_survive != 4`  
`group %in% c("Con7", "Con8")`  
`group == "Con7" & gd_of_birth == 20`  

### drop missing value
`!(is.na(gd0_weight))` 
Preferred:  
`drop_na(litters_data)` will remove any row with a missing value.  
`drop_na(litters_data, wt_increase)` will remove rows for which wt_increase is missing.  


## mutate
```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## arrange
Sort dataset on 'group' in descending order, then on `pups_born_alive` in ascending order.  
```{r}
head(arrange(litters_data, desc(group), pups_born_alive), 10)
```

## Pipes: %>% (Ctrl + Shift + M)
"and then", tibble goes in, tibble comes out  
using . as a placeholder for the result of the preceding call  
```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
```

```{r}
litters_data %>%
  lm(wt_gain ~ pups_born_alive, data = .) %>%
  broom::tidy()
```



