---
title: "Tidy_data"
author: "Jiayi SHi"
date: "2022-09-27"
output: github_document
---

## Join
inner joins (intersection)   
Outer joins:left join, right join, full join (union)  

## Key Functions
For tidying table:  
`pivot_longer`  
`pivot_wider`  
`separate`  
For combining multiple tables:  
`bind_rows`    
`*_join`  

```{r message=FALSE}
library(tidyverse)

options(tibble.print_min=5)
```

## pivot_longer
The package::function syntax lets you use a function from a package without loading the whole library.
```{r message=FALSE}
pulse_df = haven::read_sas('data/public_pulse_data.sas7bdat') %>% 
  janitor::clean_names()

```

```{r}
pulse_data = 
  pivot_longer(
    pulse_df,
    4:7,#c(4,5,6,7),
    names_to = 'visit', #create a variable named by names_to
    names_prefix = 'bdi_score_',  
    values_to = 'bdi'
  ) %>% 
  mutate(
    visit = replace(visit,visit == 'bl', '00m'), #replace(x,list,values)x: Name of vector; list: Elements to replace; values: Replacement values
    visit = factor(visit)
#data type: factor, has certain ordering
  ) %>% 
  arrange(id,visit)
```

### Assessment
* recode
```{r message=FALSE}
litters = read_csv('data/FAS_litters.csv') %>% 
  janitor::clean_names() %>% 
  select(litter_number,ends_with('weight')) %>% 
  pivot_longer(
    2:3,
    names_to = 'gd',
    values_to = 'weight'
  ) %>% 
  mutate(
    gd = recode(gd, 'gd0_weight' = 0,'gd18_weight'= 18)
  )#recode(.x, old=new) .x is a vector to modify
```

* if_else
```{r message=FALSE}
litters = read_csv('data/FAS_litters.csv') %>% 
  janitor::clean_names() %>% 
  select(litter_number,ends_with('weight')) %>% 
  pivot_longer(
    2:3, # or:gd0_weight:gd18_weight (select columns)
    names_to = 'gd', # old_col_names goes to new variable named 'gd'
    values_to = 'weight' # old_col_values goes to new variable named 'weight'
  ) %>% 
  mutate(
    gd = if_else(.$gd=='gd0_weight',0,18)
  )
```

* replace
```{r message=FALSE}
litters = read_csv('data/FAS_litters.csv') %>% 
  janitor::clean_names() %>% 
  select(litter_number,ends_with('weight')) %>% 
  pivot_longer(
    2:3,
    names_to = 'gd',
    values_to = 'weight'
  ) %>% 
  mutate(gd = replace(gd, gd=='gd0_weight', 0)) %>% 
  mutate(gd = replace(gd, gd=='gd18_weight', 18))
```

## pivot_wider

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result_wide = pivot_wider(
  analysis_result,
  names_from = 'time',
  values_from = 'mean'
    )
```

## bind_rows

```{r}
fellowship_king = 
  readxl::read_excel('data/LotR_Words.xlsx',range='B3:D6') %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")

lotr_tidy = bind_rows(fellowship_king,two_towers,return_king) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "words"
  ) %>% 
  mutate(race = str_to_lower(race))
```

## *_join

```{r message=FALSE}
pups = read_csv("data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    sex = recode(sex,`1`='male',`2`='female'),
    # backquote for number, regular quote for character 
  )

litters = read_csv("data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  separate(group,into = c("dose","day_of_treatment"),sep = 3) %>% 
  relocate(litter_number) %>% 
  mutate(
    wt_gain = gd18_weight-gd0_weight,
    dose = str_to_lower(dose)
  )

fas_data = left_join(pups,litters,by = "litter_number")
```

### Assessment

```{r message=FALSE}
surv_os = read_csv("data/surv_os.csv") %>% 
  setNames(c("uni","os"))
# change column names: setNames(df,c(newnamecol1,newnamecol2,...))
surv_prog = read_csv("data/surv_program_git.csv") %>% 
  setNames(c("uni","prog","exp"))

surv = full_join(surv_os,surv_prog,by = "uni")
```

Answer:  
```{r message=FALSE}
surv_os = read_csv("data/surv_os.csv") %>% 
  janitor::clean_names() %>% #不仅变小写，而且加下划线
  rename(id = what_is_your_uni, os = what_operating_system_do_you_use)

surv_pr_git = read_csv("data/surv_program_git.csv") %>% 
  janitor::clean_names() %>% 
  rename(
    id = what_is_your_uni, 
    prog = what_is_your_degree_program,
    git_exp = which_most_accurately_describes_your_experience_with_git)

left_join(surv_os, surv_pr_git)
inner_join(surv_os, surv_pr_git)
anti_join(surv_os, surv_pr_git)
# anti_join(x, y, by = NULL, copy = FALSE, ...): return all rows from x without a match in y.
anti_join(surv_pr_git, surv_os)
```

